---
title:  "2_process_ca_frp_data"
author: "JJWalker"
date: "2021-08-19"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---

####_2_process_ca_frp_data.Rmd_

This script uses the consolidated file output from "intersect_frp_data_with_fire_perimeters.py", 
which tags MODIS point with associated fire history information. 

Workflow:  
   -  Standardize column names and point ID  
   -  Rescale FRP data  
   -  Convert satellite info -> NA if fire year != MODIS year  
   -  Delete point if MODIS year preceeds fire year  
   -  Remove duplicate fire history info at each point  
   -  Merge with original datasets that have EVT info but not fire history  

_Input file_:  maxFRP_CA_gte4sqkm_yyyytoyyyy_processed_slim_plus_fire_info.csv

_Output file_:  data_from_script2.Rdata


```{r, echo = FALSE, message = FALSE}

# ---------------------------------------------------------------- #
#### Set up directories and filenames ####
# ---------------------------------------------------------------- #
load_libraries <- function(libs){
  for(lib in libs){
    if(!require(lib, character.only = TRUE)){    # Require returns TRUE if package loaded
      install.packages(lib, dependencies = TRUE) # If package couldn't load, re-install
      require(lib, character.only = TRUE)        # Load package after installing
    }
  }
}

load_libraries(c("dplyr"))

# Set paths and files
path_in <- "/home/jovyan/ca_frp"
path_data <- file.path(path_in, "data")
data_in <- 'maxFRP_CA_gte4sqkm_2001to2020_processed_slim_plus_fire_info.csv'
data_out <- "data_from_script2.RData"

```


```{r, echo = FALSE, message = FALSE}
# ---------------------------------------------------------------- #
#### Load in file of MODIS points that have associated fire data ####
# ---------------------------------------------------------------- #

# Get the relevant filename
  df_fire <- read.csv(file.path(path_data, data_in), header = T)

```



```{r, echo = FALSE, messsage = FALSE}

# Point ID ('index') is unique to each point, though the same ID can occur in multiple rows since
# a single point may have burned in multiple fires. The input format is unwieldy; replace with 
# a simpler sequence of sequential numbers.

 df_fire1 <- df_fire %>% 
            mutate(index_old = as.factor(index)) %>% 
            mutate(index = as.factor(index))

  levels(df_fire1$index) <- c(seq(1:length(levels(df_fire1$index))))
  
# Rename columns  
  oldnames = c("index", "MaxFRP", "newdate", "ALARM_DATE", "YEAR_", "FID_firep2", "FIRE_NAME", "GIS_ACRES", "latitude", "longitude", "pixelarea")
  newnames = c("ptid", "frp", "modis_date", "fire_date", "fire_year", "fireid", "fire_name", "fire_ac", "lat", "lon", "pixelarea")
  
  df_fire1 <-df_fire1 %>% 
            rename_at(vars(all_of(oldnames)), ~ newnames)

# Ensure point id is a factor
  df_fire1$ptid <- factor(df_fire1$ptid)

# Format date columns and remove old ones
  df_fire1$modis_date <- as.Date(df_fire1$modis_date, format = "%m/%d/%Y %H:%M:%S") 
  df_fire1$fire_date <- as.Date(df_fire1$fire_date, format = "%m/%d/%Y %H:%M:%S")
  df_fire1$modis_year <- year(df_fire1$modis_date)

# Incoming MODIS frp is scaled by 10; rescale
  df_fire1$frp <- 0.1 * df_fire1$frp
  
# Remove NA-only columns
  df_fire1 <- Filter(function(y) !all(is.na(y)), df_fire1)
  
# Convert satellite info -> NA if fire year != MODIS year. Delete point if MODIS year preceeds fire year.
  df_fire2 <- df_fire1 %>% 
    mutate(frp = replace(frp, fire_year != modis_year, NA),
           satellite = replace(satellite, fire_year != modis_year, NA),
           sample = replace(sample, fire_year != modis_year, NA)) %>% 
    filter(., fire_year <= modis_year)  # 
  
# Arrange columns 
  df_fire2 <- df_fire2 %>% 
    relocate(modis_year, .after = fire_year) %>% 
    relocate(modis_date, .after = fire_date)

# Since each MODIS reading is associated with all fires at that location, multiple fire histories will exist at points
# with > 1 burn. Remove duplicates by taking distinct entries associated with each point.
 
  df_fire3 <- df_fire2 %>% group_by(ptid) %>% 
      arrange(desc(modis_year)) %>% 
      distinct(fireid, fire_year, .keep_all = T)
  
```


```{r, echo = FALSE, message = FALSE, warning = FALSE}

# ---------------------------------------------------------------- #
#### Combine files that have EVT class info but no fire history ####
# ---------------------------------------------------------------- #

# Get relevant filenames
  filenames_evt <- list.files(file.path(path_data, "gee"), pattern = "*_processed.csv")

# Read all files to a single file. Suppress warnings b/c of spurious "incomplete last line" error
  suppressWarnings(files_w_evt_info <- do.call(rbind, lapply(file.path(path_data, "gee", filenames_evt), read.csv, row.names=NULL)))
  
# Remove NA-only columns
  files_w_evt_info <- Filter(function(y) !all(is.na(y)), files_w_evt_info)

# Remove duplicates; set class types to numeric; remove any rows in which class is only NA
  files_w_evt_info <- files_w_evt_info %>% 
      distinct(index, .keep_all = TRUE) %>% 
      mutate(across(starts_with("class"), as.numeric)) %>%  
      filter(across(starts_with("class"), all_vars(!is.na(.))))
  
```


```{r, echo = FALSE, message = FALSE}

# ---------------------------------------------------------------- #
#### Join fire history file and EVT file ####
# ---------------------------------------------------------------- #

# Take index and all class columns from EVT file
  df_evt_and_fire <- df_fire3 %>% left_join(select(files_w_evt_info, index, starts_with("class")), by = c("index_old" = "index"))

```


```{r, echo = FALSE, message = FALSE}

# Save data and environment settings   
  print(paste0("R data file saved to ", file.path(path_data, data_out)))
  save(df_evt_and_fire, file = file.path(path_data, data_out))

```

