---
title:  "Calculate burn info"
author: "JJWalker"
date: "2021-08-25"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---


####_4_calculate_burn_info.Rmd_

This script ingests the full df produced in 3_create_new_fire_df.Rmd and gnerates the burn information (burn number,
years between burns, etc).

_Input files_:  data_from_script3.RData

_Output file_:  data_from_script4.RData


```{r, echo=FALSE, message=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
#### Set up directories and filenames ####
# ---------------------------------------------------------------- #

load_libraries <- function(libs){
  for(lib in libs){
    if(!require(lib, character.only = TRUE)){    # Require returns TRUE if package loaded
      install.packages(lib, dependencies = TRUE) # If package couldn't load, re-install
      require(lib, character.only = TRUE)        # Load package after installing
    }
  }
}
#  Try/install packages
load_libraries(c("lubridate", "dplyr")) # "kableExtra", "knitr") )
#devtools::install_github("haozhu233/kableExtra")
#library(kableExtra)

# Set paths and files
path_in <-  "/home/jovyan/ca_frp"
path_data <- file.path(path_in, "data")
data_in <- "data_from_script3.RData"
data_out <- "data_from_script4.RData"
path_r <- file.path(path_in, 'R')
source(file.path(path_r, "ca_functions.R"))

```

```{r, echo = FALSE}

load(file.path(path_data, data_in))

```


```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Define EVT columns ####
# ---------------------------------------------------------------- #
# Read in data file from GDrive:  "MyDrive/Work/projects/ca_frp/evt_2014_classes_ca.csv"

# data_url <- "1y3DCmYOZFBf0G2_VfdnMHFHAyo2pfGpW" # google file ID for 2014 EVT classes
# data_url <-  "1S9AHv-NC5Jxz5NC8BxO2vU-RcV2f0hHA"  # google ID for 2001 EVT classes

# 2001 Sleeter grouped classes
data_url <-  "1l3-MmB6ut3GAUyziFyryW_yw4YmwtJUH"  # google ID for 2001 EVT classes
evt_classes <- read.csv(sprintf("https://docs.google.com/uc?id=%s&export=download", data_url))

# Change column names to conform to existing 
#names(evt_classes) <- c('evt_number', 'description', 'evt_group') # headings for 2014 data
#names(evt_classes) <- c('evt_number', 'description', 'evt_group', 'evt_group2', 'evt_fuel', 'link') # 2001 
names(evt_classes) <- c('evt_group_desc', 'evt_group', 'evt_desc', 'evt_number', 'color', 'legend', 'description', 'isautoname')  

evt_classes <- evt_classes %>% 
               select(-c(color:isautoname))

```


```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Function - Calculate burn number ####
# ---------------------------------------------------------------- #
# Count the number of times each point appears: 1 = 1st fire, 2 = 2nd, etc.
  get_fire_order = function(df) {
    df <- df %>% 
      group_by(ptid) %>% 
      arrange(fire_year) %>%  # put in ascending order
      mutate(burn_num = seq_along(fire_year)) # assign a sequence number
    df <- data.frame(df)
    df$reburn <- 0 
    
    df <- df %>%
       mutate(reburn = ifelse(burn_num > 1, 1, 0)) # reburn: no = 0, yes = 1
    df$year_int <- ave(df$fire_year, factor(df$ptid), FUN=function(t) c(0, diff(t))) # get years between fires
  return(df)
  }

```


```{r, echo = FALSE}

# Remove all points that have an FRP reading wthout an associated fire ID
# this step constrains points to only those within identified fire boundaries

tmp <- subset(x_fires_noevt, is.na(fireid) & !is.na(frp))
x_fires_noevt <- x_fires_noevt %>% filter(!(ptid %in% tmp$ptid))
                                          
```

```{r}

# Remove all points in scan angles > 27degrees (only samples 334 - 1020)

#x_fires_noevt <- x_fires_noevt %>% filter(sample >= 334 & sample <= 1020)

```


```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Get burn number for each point ####
# ---------------------------------------------------------------- #
x_fires_noevt <- get_fire_order(x_fires_noevt)

```


```{r, echo = FALSE}

# ---------------------------------------------------------------- #
#### Merge fire df with EVT classes ####
# ---------------------------------------------------------------- #
# Necessary housekeeping
x_fires_noevt$ptid <- as.factor(x_fires_noevt$ptid)
x_fires_noevt$satellite <- as.factor(x_fires_noevt$satellite)
x_classes$ptid <- as.numeric(x_classes$ptid)

# Merge
x_fires_evt <- merge(x_fires_noevt, x_classes, by = c("ptid"), all.x = T)

```
The ingested dataset has redundancies because FRP pixels are paired with each coincident fire perimeter from all previous years. After processing, an example point id has this format:

`r x_fires_noevt %>%
  filter(ptid == x_fires_evt[x_fires_evt$burn_num == 4,][[1]][1])# %>% 
#  kable() %>%
#  kable_styling()`
  
Fire year is assumed to be MODIS year for all fires 2002 - 2016. Other notes:

  * FRP = NA if the fire was prior to the MODIS era (documented only through the BLM history database)  
  * FRP pixels without a corresponding BLM database fire have BLM fire ID = NA  
  * Points with an associated BLM fire ID after 2001 but without a corresponding FRP are excluded; although the fire could've been missed due to clouds, it could also have been that the area didn't burn (the GIS database does not differentiate between burned and unburned areas within fire perimeters)
  * Satellite 0 = Terra, 1 = Aqua

###EVT classes

After EVT classes are added back to the dataframe, the process is to:

1. Determine EVT class with the greatest proportion in each 1-km MODIS pixel
2. Calculate area of that EVT class in each pixel
3. Remove pixels in which none of the classes had a majority

```{r, echo = FALSE, warning = FALSE}

# ---------------------------------------------------------------- #
#### Process EVT classes ####
# ---------------------------------------------------------------- #
  
  # Record the EVT class that had the greatest area in each 1-km MODIS pixel
  x_fires_evt$max_evt_cat <- apply(x_fires_evt, 1, function(y) names(y[grepl("class", names(y))])[which.max(y[grepl("class", names(y))])[1]])
  
  # ...record the area that class represented
  x_fires_evt$max_evt_area <- apply(x_fires_evt[grepl("class", names(x_fires_evt))], 1, max, na.rm = TRUE)
  
  # ...calculate the proportion of pixels that class represented
  x_fires_evt$max_evt_prop <- x_fires_evt$max_evt_area/x_fires_evt$evt_sum
  
  # Remove rows in which max_class == NA; this means none of the classes had a majority
  x_fires_evt <- x_fires_evt[!is.na(x_fires_evt$max_evt_cat), ]
  
  # Convert to factor
  x_fires_evt$max_evt_cat <- as.factor(x_fires_evt$max_evt_cat)
  
  # Trim "class" prefix off classes
  x_fires_evt$max_evt_cat <- gsub('class', '', x_fires_evt$max_evt_cat)
  
  x_fires_evt <- x_fires_evt %>% 
                 mutate(max_evt_cat = as.numeric(max_evt_cat))
  
  # Merge with file of EVT classes to associate a name with each class
  x_fires_evt <- left_join(x_fires_evt, evt_classes, c("max_evt_cat" = "evt_number"))
  
  # Get rid of all individual classes. dplyr rocks!
  x_fires_evt <- x_fires_evt %>% select(-starts_with('class'))
  
  # Convert burn_num to factor
  x_fires_evt$burn_num <- as.factor(x_fires_evt$burn_num)
  
  # Rearrange 
#  x_fires_evt <- x_fires_evt %>% select(ptid:evt_number, max_evt_cat)

```

`r x_fires_evt %>%
  filter(ptid == 5767) %>% 
  arrange(fire_year) %>% 
  kable() %>%
  kable_styling()`




####Save data

RData file is saved to `r file.path(path_r, data_out)`
```{r, echo=FALSE}

# Save data and environment settings   
print(paste0("R data file saved to ", file.path(path_data, data_out)))
save.image(file = file.path(path_data, data_out))

```