---
title:  "CA FRP bowtie exploration"
author: "JJWalker"
date: "2021-08-25"
output:
  html_notebook:
    highlight: tango
    mathjax: null
    number_sections: no
    theme: spacelab
---


####_CA FRP bowtie exploration.Rmd

This script calculates FRP adjustments  

_Input files_:  data_from_script4_k.RData  

_Output file_:  data_from_bowtie_correction.RData


Here we use the file x_fires_evt to explore how to best correct the effects of look angle on FRP.  
This file relies on ALL data point--snot the subset based on sample. 

```{r, echo=FALSE, message=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
#### Set up directories and filenames ####
# ---------------------------------------------------------------- #

load_libraries <- function(libs){
  for(lib in libs){
    if(!require(lib, character.only = TRUE)){    # Require returns TRUE if package loaded
      install.packages(lib, dependencies = TRUE) # If package couldn't load, re-install
      require(lib, character.only = TRUE)        # Load package after installing
    }
  }
}
#  Try/install packages
load_libraries(c("lubridate", "dplyr", "kableExtra", "knitr"))
#devtools::install_github("haozhu233/kableExtra")

# Set paths
path_in = 'E:/OneDrive - DOI/projects/ca_frp'
path_data <- file.path(path_in, "data")
data_in <- "data_from_script4_k.Rdata"
path_r <- file.path(path_in, 'R')
data_out <- "data_adjusted_by_sample.Rdata"
source(file.path(path_r, "ca_functions.R"))

```

  
```{r, echo = F, warning = F}

# ---------------------------------------------------------------- #
#### Correct for bowtie MODIS effect ####
# ---------------------------------------------------------------- #

# Calculate mean and median by sample for combined Terra/Aqua

x_bysample_all <- x_fires_evt %>% 
  group_by(sample, satellite) %>% 
  summarize(frp_mean = mean(frp), frp_med = median(frp))

```
 
Terra and Aqua distributions are not identical, so calculate polynomial fits separately for each.
 
```{r, echo=FALSE}

# Subset main dataset to Terra or Aqua fires only
x_fires_evt.t <- subset(x_fires_evt, satellite == '0')
x_fires_evt.a <- subset(x_fires_evt, satellite == '1')

```
 
 
 Raw FRP values by sample
 
```{r,echo = FALSE, warning=FALSE}

ggplot(x_fires_evt, aes(sample, frp, group = satellite, color = satellite)) + geom_point() +
  scale_y_log10() +
  scale_color_hue(labels = c('Terra', 'Aqua'))

```
Simply remove any points outside of 27º (Heward et al, 2013). That's 1/2 of the range: 1354/2 = 667  ~= 334 - (1354 - 334 = 1020)

```{r, echo = FALSE}

# Remove all points in scan angles > 27degrees (only samples 334 - 1020)

# x_fires_noevt <- x_fires_noevt %>% filter(sample >= 334 & sample <= 1020)

```


Manually calculate line for Terra points
```{r}
# t - manually calculate line

points <- data.frame('x' = c(200, 250, 300, 400, 500, 600, 700, 800, 900, 1000, 1050, 1100,
                             1150), 
                     'y'= c(3900, 3100, 2590, 2100, 1800, 1680, 1580, 1710, 1910, 2380, 2620, 3100,
                            3750))

ggplot(x_fires_evt.t, aes(sample, frp)) + geom_point() + 
  scale_x_continuous(breaks = seq(0, 1354, by = 100)) + 
  scale_y_continuous(breaks = seq(0, 10000, by = 500)) + 
  geom_line(data = points, aes(x,y), color = 'red')


```
Try to get the polynomial equation for the line
```{r, echo = F}

# get equation
fit_t.tp <- lm(y ~ poly(x, 2, raw = TRUE), data = points)
coeff0 <- summary(fit_t.tp)$coefficients[1, 1] 
coeff1 <- summary(fit_t.tp)$coefficients[2, 1] 
coeff2 <- summary(fit_t.tp)$coefficients[3, 1]

points$fit <- coeff0 + (coeff1 * points$x) + (coeff2 * (points$x^2))

tmp <- data.frame('pt' = seq(0, 1353))
tmp$pt_fit <- coeff0 + (coeff1 * tmp$pt) + (coeff2 * (tmp$pt^2))

ggplot(x_fires_evt.t, aes(sample, frp)) + geom_point() + 
  scale_x_continuous(breaks = seq(0, 1354, by = 100)) + 
  scale_y_continuous(breaks = seq(0, 10000, by = 500)) + 
  geom_line(data = points, aes(x, fit), color = 'red') +
  geom_line(data = tmp, aes(pt, pt_fit), color = "blue")

```
It just doesn't correct the data properly
```{r}

frp.min <- min(points$fit, na.rm = T)

x_fires_evt.t$fit_all <- coeff0 + (coeff1 * x_fires_evt.t$sample) + (coeff2 * (x_fires_evt.t$sample^2))

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x_fires_evt.t$adj_all <- frp.min/x_fires_evt.t$fit_all

# Multiply the actual frp by that adjustment 
x_fires_evt.t$frp_adj_all <- x_fires_evt.t$frp * x_fires_evt.t$adj_all

ggplot(x_fires_evt.t, aes(sample, frp_adj_all)) + geom_point(color = 'blue') +
  geom_point(aes(sample, fit_all), color = "red") 

```

 TERRA - Calculate polynomial fit for ALL FRP values



```{r, echo=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
####  TERRA - Calculate polynomial fit for ALL FRP values ####
# ---------------------------------------------------------------- #

fit_all <- lm(frp ~ poly(sample, 2, raw = TRUE), data = x_fires_evt.t)
coeff0 <- summary(fit_all)$coefficients[1, 1]
coeff1 <- summary(fit_all)$coefficients[2, 1]
coeff2 <- summary(fit_all)$coefficients[3, 1]

x_fires_evt.t$fit_all <- coeff0 + (coeff1 * x_fires_evt.t$sample) + (coeff2 * (x_fires_evt.t$sample^2))

# Added
x_fires_evt.t$test <- 27546 + (22.17 * x_fires_evt.t$sample) + (0.0095 * (x_fires_evt.t$sample^2))

# Adjust data  
frp.min <- min(x_fires_evt.t$fit_all, na.rm = T)

# Added
frp.mint <- min(x_fires_evt.t$test, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x_fires_evt.t$adj_all <- frp.min/x_fires_evt.t$fit_all

x_fires_evt.t$adj_test <- frp.mint/x_fires_evt.t$test
x_fires_evt.t$frp_adj_test <- x_fires_evt.t$frp + x_fires_evt.t$adj_test

# Multiply the actual frp by that adjustment 
x_fires_evt.t$frp_adj_all <- x_fires_evt.t$frp * x_fires_evt.t$adj_all

```

```{r,echo = FALSE, warning=FALSE}

ggplot(x_fires_evt.t, aes(sample, frp)) + geom_point() +
  scale_y_log10() 

```

Fit a polynomial correction line to all Terra FRP values:

```{r, echo= FALSE, warning= FALSE, message = FALSE}

# ---------------------------------------------------------------- #
#### TERRA - Plot adjustment line across samples ####
# ---------------------------------------------------------------- #

ggplot(x_fires_evt.t, aes(sample, frp)) + geom_point() +
  geom_point(aes(sample, fit_all), color = "red") +
  scale_y_log10()

```

When applied to the raw FRP values, the correction looks ok but not great. The lack of lower values at the extreme look angles is likely due to the decreased sensitivity of the sensor to lower FRP values there.

```{r, echo= FALSE, warning= FALSE}
# ---------------------------------------------------------------- #
#### TERRA - Plot all FRP points by sample, with adjusted data ####
# ---------------------------------------------------------------- #

ggplot(x_fires_evt.t, aes(sample, frp)) + 
  geom_point() +
  geom_point(data = x_fires_evt.t, aes(sample, frp_adj_all), color = "blue") +
  scale_y_log10()
```


```{r, echo=F, warning=FALSE}

# ---------------------------------------------------------------- #
# TERRA - Calculate polynomial fit on FRP values by sample ####
# ---------------------------------------------------------------- #

# Summarize by sample
x.bysample.t <- x_fires_evt.t %>% 
  group_by(sample) %>% 
  summarize(frp_mean = mean(frp), frp_med = median(frp))

# Calculate polynomial fit - MEDIAN
fit_med <- lm(frp_med ~ poly(sample, 2, raw = TRUE), data = x.bysample.t)
coeff0 <- summary(fit_med)$coefficients[1, 1]
coeff1 <- summary(fit_med)$coefficients[2, 1]
coeff2 <- summary(fit_med)$coefficients[3, 1]
x.bysample.t$fit_med <- coeff0 + (coeff1 * x.bysample.t$sample) + (coeff2 * (x.bysample.t$sample^2))

# Calculate polynomial fit - MEAN
fit_mean <- lm(frp_mean ~ poly(sample, 2, raw = TRUE), data = x.bysample.t)
coeff0 <- summary(fit_mean)$coefficients[1, 1]
coeff1 <- summary(fit_mean)$coefficients[2, 1]
coeff2 <- summary(fit_mean)$coefficients[3, 1]
x.bysample.t$fit_mean <- coeff0 + (coeff1 * x.bysample.t$sample) + (coeff2 *(x.bysample.t$sample^2))
```


```{r, echo = F, warning = FALSE}

# ---------------------------------------------------------------- #
# TERRA - Calculate and apply median and mean FRP correction
# ---------------------------------------------------------------- #

# Get minimum value of MEDIAN fit
# ---------------------------------
frp.min <- min(x.bysample.t$fit_med, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x.bysample.t$adj_med <- frp.min / x.bysample.t$fit_med

# Apply the correction in the consolidated database
x.bysample.t$frp_adj_med <- x.bysample.t$frp_med * x.bysample.t$adj_med

# Associate the sample # in the full database with the adjustment ratio
x_fires_evt.t <- left_join(x_fires_evt.t, distinct(dplyr::select(x.bysample.t, sample, adj_med)), by = "sample")

# Multiply the actual frp by that adjustment 
x_fires_evt.t$frp_adj_med <- x_fires_evt.t$frp * x_fires_evt.t$adj_med


# Get minimum value of MEAN fit
# ---------------------------------
frp.min <- min(x.bysample.t$fit_mean, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x.bysample.t$adj_mean <- frp.min / x.bysample.t$fit_mean

# Apply the correction in the consolidated database
x.bysample.t$frp_adj_mean <- x.bysample.t$frp_mean * x.bysample.t$adj_mean

# Associate the sample # in the full database with the adjustment ratio
x_fires_evt.t <- left_join(x_fires_evt.t, distinct(dplyr::select(x.bysample.t, sample, adj_mean)), by = "sample")

# Multiply the actual frp by that adjustment 
x_fires_evt.t$frp_adj_mean <- x_fires_evt.t$frp * x_fires_evt.t$adj_mean

```

#### Terra MEDIAN with polynomial line fit

```{r, echo=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
####  Plot TERRA MEDIAN ###
# ---------------------------------------------------------------- #

ggplot(x.bysample.t, aes(sample, frp_med)) + 
  geom_point() + 
  geom_line(data = x.bysample.t, aes(sample, fit_med, color = satellite), size = 1.2, color = "red") +
  theme_bw() +
  labs(x = "Sample", y = "Median FRP") 
```

#### Terra MEDIAN with polynomial line fit and adjusted points

```{r, echo=F, warning=FALSE}

# ---------------------------------------------------------------- #
#### Look at improvement in median for TERRA ####
# ---------------------------------------------------------------- #

ggplot(x.bysample.t, aes(sample, frp_med)) + 
  geom_point() + 
  geom_line(aes(sample, fit_med, color = satellite), size = 1.2, color = "red") +
  geom_point(aes(sample, frp_adj_med), color = 'blue') +
  theme_bw() +
  labs(x = "Sample", y = "Median FRP") 
```

#### Terra MEAN with polynomial line fit

```{r, echo=FALSE, warning=FALSE}
# ---------------------------------------------------------------- #
#### Plot TERRA MEAN ####
# ---------------------------------------------------------------- #

ggplot(x.bysample.t, aes(sample, frp_mean)) + 
  geom_point() + 
  geom_line(aes(sample, fit_mean, color = satellite), size = 1.2, color = "red") +
  theme_bw() +
  labs(x = "Sample", y = "Mean FRP") # +
 # scale_y_log10()
```
#### Terra MEAN with polynomial line fit and adjusted points

```{r, echo=F, warning=FALSE}

# ---------------------------------------------------------------- #
#### Look at improvement in mean for TERRA ####
# ---------------------------------------------------------------- #

ggplot(x.bysample.t, aes(sample, frp_mean)) + 
  geom_point() + 
  geom_line(aes(sample, fit_mean, color = satellite), size = 1.2, color = "red") +
  geom_point(aes(sample, frp_adj_mean), color = 'blue') +
  theme_bw() +
  labs(x = "Sample", y = "Mean FRP") # +
  #scale_y_log10()

```


```{r, echo=F, warning=FALSE}

# ---------------------------------------------------------------- #
#### AQUA - Calculate polynomial fit on FRP values by sample ####
# ---------------------------------------------------------------- #

# Summarize by sample
x.bysample.a <- x_fires_evt.a %>% 
  group_by(sample) %>% 
  summarize(frp_mean = mean(frp), frp_med = median(frp))

# Calculate polynomial fit - MEDIAN
fit_med <- lm(frp_med ~ poly(sample, 2, raw = TRUE), data = x.bysample.a)
coeff0 <- summary(fit_med)$coefficients[1, 1]
coeff1 <- summary(fit_med)$coefficients[2, 1]
coeff2 <- summary(fit_med)$coefficients[3, 1]
x.bysample.a$fit_med <- coeff0 + (coeff1 * x.bysample.a$sample) + (coeff2 * (x.bysample.a$sample^2))

# Calculate polynomial fit - MEAN
fit_mean <- lm(frp_mean ~ poly(sample, 2, raw = TRUE), data = x.bysample.a)
coeff0 <- summary(fit_mean)$coefficients[1, 1]
coeff1 <- summary(fit_mean)$coefficients[2, 1]
coeff2 <- summary(fit_mean)$coefficients[3, 1]
x.bysample.a$fit_mean <- coeff0 + (coeff1 * x.bysample.a$sample) + (coeff2 *(x.bysample.a$sample^2))


```


```{r, echo=FALSE}

# AQUA - manually calculate line 

points <- data.frame('x' = c(200, 250, 300, 400, 500, 600, 700, 800, 900, 1000, 1050, 1100,
                             1150), 
                     'y'= c(3900, 3200, 2790, 2100, 1800, 1680, 1580, 1710, 1910, 2380, 2620, 3100,
                            3750))

ggplot(x_fires_evt.a, aes(sample, frp)) + geom_point() + 
  scale_x_continuous(breaks = seq(0, 1354, by = 100)) + 
  scale_y_continuous(breaks = seq(0, 10000, by = 500)) + 
  geom_line(data = points, aes(x,y), color = 'red')

```
```{r}

#AQUA from Freeborn et al, 2011:  y = −0.0095x2 + 22.17x + 27546
```



```{r, echo = F, warning = FALSE}

# ---------------------------------------------------------------- #
# AQUA - Calculate and apply median and mean FRP correction
# ---------------------------------------------------------------- #

# Get minimum value of MEDIAN fit
# ---------------------------------
frp.min <- min(x.bysample.a$fit_med, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x.bysample.a$adj_med <- frp.min / x.bysample.a$fit_med

# Apply the correction in the consolidated database
x.bysample.a$frp_adj_med <- x.bysample.a$frp_med * x.bysample.a$adj_med

# Associate the sample # in the full database with the adjustment ratio
x_fires_evt.a <- left_join(x_fires_evt.a, distinct(dplyr::select(x.bysample.a, sample, adj_med)), by = "sample")

# Multiply the actual frp by that adjustment 
x_fires_evt.a$frp_adj_med <- x_fires_evt.a$frp * x_fires_evt.a$adj_med


# Get minimum value of MEAN fit
# ---------------------------------
frp.min <- min(x.bysample.a$fit_mean, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio for each sample
x.bysample.a$adj_mean <- frp.min / x.bysample.a$fit_mean

# Apply the correction in the consolidated database
x.bysample.a$frp_adj_mean <- x.bysample.a$frp_mean * x.bysample.a$adj_mean

# Associate the sample # in the full database with the adjustment ratio
x_fires_evt.a <- left_join(x_fires_evt.a, distinct(dplyr::select(x.bysample.a, sample, adj_mean)), by = "sample")

# Multiply the actual frp by that adjustment 
x_fires_evt.a$frp_adj_mean <- x_fires_evt.a$frp * x_fires_evt.a$adj_mean

```


#### Aqua MEDIAN with polynomial line fit

```{r, echo=FALSE, warning=FALSE}

# ---------------------------------------------------------------- #
####  Plot TERRA MEDIAN ###
# ---------------------------------------------------------------- #

ggplot(x.bysample.a, aes(sample, frp_med)) + 
  geom_point() + 
  geom_line(aes(sample, fit_med, color = satellite), size = 1.2, color = "red") +
  theme_bw() +
  labs(x = "Sample", y = "Median FRP") 
```


#### Aqua MEDIAN with polynomial line fit and adjusted points

```{r, echo=F, warning=FALSE}

# ---------------------------------------------------------------- #
#### Look at improvement in median for AQUA ####
# ---------------------------------------------------------------- #

ggplot(x.bysample.a, aes(sample, frp_med)) + 
  geom_point() + 
  geom_line(aes(sample, fit_med, color = satellite), size = 1.2, color = "red") +
  geom_point(aes(sample, frp_adj_med), color = 'blue') +
  theme_bw() +
  labs(x = "Sample", y = "Median FRP") 
```



```{r, echo=FALSE, warning=FALSE}

# plot of all median points (no satellite distinction) and separately calculated fits
ggplot(temp.all, aes(sample, frp_med)) + 
  geom_point() + 
  geom_line(data = temp.a, aes(sample, fit_med, color = 'aqua'), size = 1.2) +
  geom_line(data = temp.t, aes(sample, fit_med, color = 'terra'), size = 1.2) +
  theme_bw()

```



```{r, echo=F, warning=FALSE}

# Correct separately - AQUA

#Make adjustments - MEDIAN - AQUA
frp.min <- min(temp.a$fit_med, na.rm = T)

# Apply the minimum across the fit curve to calculate an adjustment ratio
temp.a$adj_med <- frp.min/temp.a$fit_med

# Now associate the sample # in the full database with the adjustment ratio
merge.a <- left_join(x_fires_evt.aqua, distinct(dplyr::select(temp.a, sample, adj_med)), by = "sample")

# Multiply the actual frp by that adjustment 
merge.a$frp_adj <- merge.a$frp * merge.a$adj_med

```


```{r, echo=F, warning=FALSE}

# Look at improvement in median for AQUA

test.a <- merge.a %>% group_by(sample) %>% summarize(frp_med = median(frp_adj))

ggplot(test.a, aes(sample, frp_med)) + geom_point(color = "blue") +
  geom_point(data=temp.a, aes(sample, frp_med), color = "black") + scale_y_log10()

```


```{r}
# combine the corrected datasets


x_fires_evt.nosat <- subset(x_fires_evt, is.na(satellite))
x_fires_evt.nosat$adj_med <- NA
x_fires_evt.nosat$frp_adj <- NA

x_fires_evt.new <- rbind(merge.a, merge.t, x_fires_evt.nosat)

```
